---
title: "Get started with bootstrict"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with bootstrict}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`bootstrict` re-implements the **Bootstrap 5.3** layout, content, forms and
component library as Shiny UI functions. Every widget mirrors the Bootstrap 5.3
HTML structure **one-to-one**, so a designer's mockup and exported SASS
variables drop straight into a Shiny application — with **minimum deviation from
Shiny itself**.

```{r}
library(shiny)
library(bootstrict)
```

## Why bootstrict

Working with a designer who does not know Shiny is awkward for two reasons:

- some Shiny components are *not* plain Bootstrap; and
- some Bootstrap components are *missing* from Shiny.

`bootstrict` closes that gap by giving you the whole Bootstrap 5 surface, and
**nothing more** — so you can tell a designer: *"use anything from
[the Bootstrap 5.3 docs](https://getbootstrap.com/docs/5.3/), but nothing else."*
Bootstrap 5.3 is vendored by the package itself and compiled with
[`sass`](https://rstudio.github.io/sass/), so the version reaching the browser
is the one `bootstrap_version()` reports.

## A first app

Every app starts with a page constructor (`bs_page()`), a theme, and Bootstrap
widgets composed like any other Shiny UI:

```{r}
ui <- bs_page(
  title = "Sign in",
  theme = bootstrict_theme(primary = "#ff6600"),
  bs_container(
    bs_card(
      bs_card_header("Sign in"),
      bs_card_body(
        bs_text_input("email", "Email", placeholder = "you@example.com"),
        bs_password_input("pw", "Password"),
        bs_button("go", "Sign in", color = "primary")
      )
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$go, {
    print(input$email)
    print(input$pw)
  })
}

shinyApp(ui, server)
```

## Conventions (minimum deviation from Shiny)

Once you know these four rules you know how *every* constructor in the package
behaves:

- **Naming.** Every constructor is `snake_case`, prefixed `bs_`
  (`bs_card()`, `bs_text_input()`, …). Nothing masks a Shiny function.
- **`...` works exactly like Shiny/htmltools.** **Named** arguments become HTML
  attributes; **unnamed** arguments become children. Extra `class` values passed
  through `...` are merged with the component's own classes.
- **Interactive widgets take a leading `id`.** Their value is then available as
  `input$id`, just like a Shiny input.
- **Form inputs delegate to the matching `shiny::*Input()`.** So the reactive
  value and every `updateXxx()` keep working identically — `bootstrict` only
  layers Bootstrap 5 markup, sizing, help text, switches, input groups and
  floating labels on top.

## The map of the package

`bootstrict` covers the whole Bootstrap 5.3 catalogue. Each area has its own
article:

| Article | What it covers |
| --- | --- |
| [Theming and the designer hand-off](theming.html) | `bootstrict_theme()`, SASS variable sheets, colour modes |
| [Layout and the grid](layout.html) | pages, containers, the 12-column grid, stacks |
| [Content: tables, media and typography](content.html) | `bs_table()`, images, figures, blockquotes, headings, lists |
| [Forms and inputs](forms.html) | every input, input groups, floating labels, validation |
| [Components](components.html) | accordion, alert, badge, buttons, card, carousel, collapse, list group, progress, spinner, placeholder |
| [Navigation](navigation.html) | nav & tabs, navbar, breadcrumb, pagination, dropdown |
| [Overlays and server-driven interactivity](overlays.html) | modal, offcanvas, toast, tooltips/popovers, the `update_bs_*()` pattern |

## How bootstrict differs from Shiny

`bootstrict` stays as close to Shiny as it can, but a handful of behaviours
differ **on purpose**, to follow native Bootstrap. If you already know Shiny,
these are the things to watch for — each is covered in depth in the article
listed.

### Overlays live in the UI — they are not built from the server

Shiny builds modals and notifications on the server
(`showModal(modalDialog(...))`). `bootstrict` follows the native Bootstrap
pattern: the modal, toast or offcanvas is declared **once in the UI** with an
`id`, and the server only **opens or closes it by id**.

| Task | Shiny | bootstrict |
| --- | --- | --- |
| Open a modal | `showModal(modalDialog(...))` | declare `bs_modal("id", …)`, then `show_bs_modal("id")` |
| Close a modal | `removeModal()` | `hide_bs_modal("id")` |
| Notification | `showNotification("…")` | `bs_notify_toast("…")` |
| Offcanvas / drawer | *(not in Shiny)* | declare `bs_offcanvas("id", …)`, then `show_bs_offcanvas("id")` |

Place overlay widgets at the **top level of the page** (a direct child of
`bs_page()` / `bs_container()`), never nested inside a `bs_card()` — Bootstrap
can otherwise clip or mis-position them. See
[Overlays](overlays.html).

### Server helpers take `id` first and `session` last

Shiny's updaters take the session first: `updateTextInput(session, "id", …)`.
Every `bootstrict` helper takes the **id first** and the **session last and
optional** (it defaults to the current reactive domain), and ids are namespaced
automatically inside modules:

```{r}
update_bs_tabset("tabs", selected = "profile") # no session argument needed
show_bs_modal("info")
```

### Two kinds of inputs

1. **Inputs that delegate to Shiny** — `bs_text_input()`, `bs_numeric_input()`,
   `bs_select_input()`, `bs_radio_input()`, `bs_checkbox_input()`,
   `bs_checkbox_group_input()`, `bs_file_input()`, `bs_textarea_input()`,
   `bs_password_input()`. They wrap the matching `shiny::*Input()`, so
   `input$id` **and Shiny's own `updateXxx()` keep working unchanged**.
2. **Native inputs** — `bs_range_input()`, `bs_color_input()`,
   `bs_date_input()`, `bs_date_range_input()` and the `.btn-check` segmented
   controls (`bs_radio_button_input()`, `bs_checkbox_button_input()`). Either
   Shiny has no equivalent, or its equivalent is not Bootstrap markup. They
   ship their own bindings, so drive them with `update_bs_range()` /
   `update_bs_color()` / `update_bs_date_input()` /
   `update_bs_toggle_buttons()`.

### `bs_button()` is an action button only when given an id

`bs_button("go", "Go")` behaves exactly like `shiny::actionButton()` —
`input$go` is the click count. Called **without** an `id` it is an inert, styled
button; reactivity is opt-in.

## Run the bundled apps

```{r}
# The catalogue: every widget, one card each.
shiny::runApp(system.file("examples/demo", package = "bootstrict"))

# The showcase: Quake Watch, a seismic monitor built on datasets::quakes.
shiny::runApp(system.file("examples/quakewatch", package = "bootstrict"))
```
