---
title: "Components"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Components}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

This article covers the Bootstrap component library: cards, alerts, badges,
buttons, accordions, carousels, collapses, list groups, progress bars, spinners
and placeholders. Navigation components (nav/tabs, navbar, breadcrumb,
pagination, dropdown) have [their own article](navigation.html); overlays
(modal, offcanvas, toast, tooltips) are in [Overlays](overlays.html).

Some of these components are **interactive** — they report state as `input$id`
and are driven from the server with an `update_bs_*()` helper. Those follow the
same pattern throughout `bootstrict`; see [Overlays and server-driven
interactivity](overlays.html) for the full story.

## Buttons

`bs_button()` renders a Bootstrap button. It becomes a Shiny action button
**only when given an `id`** — then `input$id` is the click count, exactly like
`shiny::actionButton()`. Without an `id` it is an inert, styled button.

```{r}
bs_button(
  id = NULL,
  label = NULL,
  ...,
  color = "primary",   # a theme colour or "link"
  outline = FALSE,     # .btn-outline-*
  size = NULL,         # "sm" or "lg"
  disabled = FALSE,
  href = NULL,         # set to render an <a> styled as a button
  type = "button",     # "button"/"submit"/"reset"
  class = NULL
)
```

```{r}
bs_button("go", "Go", color = "primary")
bs_button(label = "Cancel", color = "secondary", outline = TRUE)
```

Group buttons with `bs_button_group()` (add `vertical = TRUE` for a column), and
group groups with `bs_button_toolbar()`:

```{r}
bs_button_group(
  bs_button(label = "Left"),
  bs_button(label = "Right")
)
```

The standalone Bootstrap close button (used inside dismissible components) is
`bs_close_button()`. On a dark background use `white = TRUE`.

## Cards

Cards are built from composable pieces. `bs_card()` is the container; the rest
(`bs_card_header()`, `bs_card_body()`, `bs_card_title()`,
`bs_card_subtitle()`, `bs_card_text()`, `bs_card_footer()`, `bs_card_link()`)
map one-to-one to the Bootstrap card sub-elements.

```{r}
bs_card(
  bs_card_header("Featured"),
  bs_card_body(
    bs_card_title("Card title"),
    bs_card_subtitle("A subtitle"),
    bs_card_text("Some quick example text to build on the card title."),
    bs_card_link("Read more", href = "#")
  ),
  bs_card_footer("Updated 3 mins ago")
)
```

`bs_card()` accepts `color=` (a background theme colour, `.text-bg-*`) and
`border=` (a border colour). For images, `bs_card_img()` places a picture at the
`"top"`, `"bottom"` or as an `"overlay"`; pair the overlay with
`bs_card_img_overlay()` to float text on top. `bs_card_group()` snaps several
cards into an equal-width attached grid.

```{r}
bs_card_group(
  bs_card(bs_card_body("A")),
  bs_card(bs_card_body("B"))
)
```

## Alerts

`bs_alert()` is a coloured message box. `dismissible = TRUE` adds a close button
and fade-out.

```{r}
bs_alert("Well done! You read this important message.", color = "success")
bs_alert("Heads up — check your settings.", color = "warning", dismissible = TRUE)
```

Inside an alert, `bs_alert_heading()` gives a matched heading and
`bs_alert_link()` a colour-matched link:

```{r}
bs_alert(
  color = "danger",
  bs_alert_heading("Something went wrong"),
  "Read the ", bs_alert_link("full log", href = "#"), " for details."
)
```

## Badges

Small count/label pills:

```{r}
bs_badge("New", color = "success")
bs_badge("42", color = "danger", pill = TRUE)
```

## Accordion (interactive)

A vertically collapsing set of panels. The value(s) of the open panel(s) are
reported as `input$id`, and you open/close them from the server with
`update_bs_accordion()`.

```{r}
bs_accordion(
  "acc",
  bs_accordion_panel("First", "Panel one body", value = "one"),
  bs_accordion_panel("Second", "Panel two body", value = "two"),
  open = "one"
)
```

- `open` accepts panel value(s), or `TRUE` (open all — only sensible with
  `multiple = TRUE`), or `FALSE`/`NULL` (none).
- `multiple = TRUE` lets panels stay open independently ("always open").
- `flush = TRUE` renders edge-to-edge.

```{r}
server <- function(input, output, session) {
  observe(print(input$acc))                          # value(s) of open panel(s)
  observeEvent(input$go, update_bs_accordion("acc", open = "two"))
}
```

## Carousel (interactive)

A slideshow of `bs_carousel_item()`s. The 0-based index of the active slide is
reported as `input$id`; drive it with `update_bs_carousel()`.

```{r}
bs_carousel(
  "demo",
  bs_carousel_item(bs_img("1.jpg"), active = TRUE, caption = "First slide"),
  bs_carousel_item(bs_img("2.jpg"))
)
```

- `indicators` / `controls` toggle the clickable dots and prev/next arrows.
- `fade = TRUE` crossfades; `autoplay = TRUE` (default) auto-cycles;
  `interval` sets the cycle time in ms.
- Exactly one item is `active` (the first is defaulted if you set none).

```{r}
update_bs_carousel("demo", to = 2)       # jump to a slide (0-based)
update_bs_carousel("demo", slide = "next")
```

## Collapse (interactive)

A container that shows/hides content. Its state is reported as `input$id`
(`TRUE` when visible). Toggle it from the UI with `bs_collapse_trigger()` (no
server round trip) or from the server with `update_bs_collapse()`.

```{r}
tagList(
  bs_collapse_trigger("more", "Toggle details"),
  bs_collapse("more", "Hidden content revealed on toggle.")
)
```

```{r}
update_bs_collapse("more", "show")   # "toggle" (default), "show" or "hide"
```

## List group (interactive)

A flexible list that becomes **server-selectable** when you give it an `id`:
the active item's `value` is reported as `input$id`, and
`update_bs_list_group()` activates an item by value.

```{r}
bs_list_group(
  "picker",
  bs_list_group_item("An item", value = "a", action = TRUE),
  bs_list_group_item("A second item", value = "b", action = TRUE)
)
```

Items can be plain (`<li>`), actionable (`action = TRUE` → a `<button>`), or
links (`href=` → an `<a>`). `flush = TRUE`, `numbered = TRUE` and
`horizontal = TRUE` change the layout; `color=` gives an item a contextual
colour.

```{r}
update_bs_list_group("picker", selected = "b")
```

## Progress (interactive)

`bs_progress()` is the track; `bs_progress_bar()` is a filled bar. Pass **two or
more** bars to `bs_progress()` and it renders a Bootstrap 5.3
`.progress-stacked` group. Give a bar an `id` to update it from the server with
`update_bs_progress()`.

```{r}
bs_progress(bs_progress_bar(value = 25, id = "load", color = "success"))

# stacked (5.3):
bs_progress(
  height = "1.5rem",
  bs_progress_bar(value = 15, color = "success", label = "15%"),
  bs_progress_bar(value = 30, color = "danger",  label = "30%")
)
```

```{r}
update_bs_progress("load", value = 80, label = "80%")
```

A bar supports `striped`, `animated` (implies striped), `min`/`max` bounds, and
an `aria_label` for accessibility.

## Spinners and placeholders

`bs_spinner()` is a loading indicator — a spinning border or a pulsing dot:

```{r}
bs_spinner(type = "border", color = "primary")
bs_spinner(type = "grow", color = "secondary", size = "sm")
```

Placeholders are "skeleton" blocks shown while content loads. `bs_placeholder()`
is one block (`width` is a 1–12 column count); wrap the whole skeleton in
`bs_placeholder_glow()` or `bs_placeholder_wave()` to animate it:

```{r}
bs_placeholder_glow(
  bs_placeholder(width = 7),
  bs_placeholder(width = 4),
  bs_placeholder(width = 6)
)
```
