---
title: "Content: tables, media and typography"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Content: tables, media and typography}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

These are the static "content" widgets: tables, images and figures, blockquotes,
display typography, and lists — plus a few small helpers. None report state to
the server; they are pure UI.

## Tables

`bs_table()` renders a faithful Bootstrap `<table class="table">`. Pass a data
frame or matrix as the first argument and it builds the header and body for you;
pass tag children instead to hand-build the rows.

```{r}
bs_table(
  data = NULL,
  ...,                 # manual children when `data` is NULL
  striped = FALSE,     # zebra-striped rows
  bordered = FALSE,    # borders on all sides
  borderless = FALSE,  # remove all borders
  hover = FALSE,       # hover highlight
  small = FALSE,       # condensed padding
  variant = NULL,      # a theme colour: .table-{colour}
  responsive = FALSE,  # TRUE, or a breakpoint "sm"/"md"/… for horizontal scroll
  align = NULL,        # vertical alignment: "middle"/"top"/"bottom"
  caption = NULL,
  class = NULL
)
```

The most common form — a styled data-frame table:

```{r}
bs_table(head(mtcars), striped = TRUE, hover = TRUE)
```

`responsive = TRUE` wraps the table so it scrolls horizontally on narrow
screens (or pass a breakpoint to only do so below it):

```{r}
bs_table(
  df,
  striped = TRUE,
  small = TRUE,
  responsive = TRUE,
  caption = "24 matching events"
)
```

## Images

`bs_img()` is a Bootstrap `<img>` with the usual responsive utilities.

```{r}
bs_img(
  src,
  ...,
  fluid = FALSE,      # .img-fluid — scale down to fit the parent
  thumbnail = FALSE,  # .img-thumbnail — rounded bordered frame
  rounded = FALSE,    # rounded corners
  object_fit = NULL,  # "contain"/"cover"/"fill"/"scale"/"none" (5.3)
  alt = NULL,
  class = NULL
)
```

```{r}
bs_img("logo.png", alt = "Logo", fluid = TRUE)
```

## Figures

A figure groups an image with a caption. Use `bs_figure_img()` and
`bs_figure_caption()` inside `bs_figure()`:

```{r}
bs_figure(
  bs_figure_img("photo.jpg", alt = "A photo"),
  bs_figure_caption("A caption for the image.")
)
```

## Blockquotes

`bs_blockquote()` renders a Bootstrap `.blockquote` inside a `<figure>`, with an
optional attribution `footer`:

```{r}
bs_blockquote(
  "The data set gives the locations of 1000 seismic events of MB > 4.0.",
  footer = "R datasets documentation"
)
```

## Display typography

Two constructors cover Bootstrap's opinionated typographic accents.

`bs_display_heading()` is a large, prominent heading; `level` (1–6) sets both
the size (`.display-{level}`) and the heading element:

```{r}
bs_display_heading("Seismic activity, Fiji region", level = 2)
```

`bs_lead()` is a standout opening paragraph (`.lead`):

```{r}
bs_lead("1,000 events of magnitude 4.0 and above, recorded since 1964.")
```

## Lists

`bs_list_unstyled()` strips the default list styling; each unnamed argument
becomes an `<li>`. `bs_list_inline()` lays the items out horizontally.

```{r}
bs_list_unstyled("First", "Second", "Third")
bs_list_inline("One", "Two", "Three")
```

You can also pass explicit `tags$li()` children when an item needs richer
content:

```{r}
bs_list_unstyled(
  tags$li(tags$code("bootstrict_theme()"), " — the designer's sheet is the theme"),
  tags$li(tags$code("bs_offcanvas()"), " — the filter drawer")
)
```

(For an interactive, selectable list, use `bs_list_group()` — see
[Components](components.html).)

## Small helpers

A handful of one-off utilities round out the content toolkit:

- **`bs_icon_link()`** — the Bootstrap 5.3 icon-link: an inline-flex `<a>`
  pairing an icon with text. `hover = TRUE` nudges the icon on hover.

  ```{r}
  bs_icon_link("Documentation", href = "#")
  ```

- **`bs_ratio()`** — wraps an embedded element (an `<iframe>`, a video) so it
  keeps a fixed aspect ratio. `ratio` is `"1x1"`, `"4x3"`, `"16x9"` or
  `"21x9"`.

  ```{r}
  bs_ratio(tags$iframe(src = "https://example.com"), ratio = "16x9")
  ```

- **`bs_vr()`** — a vertical divider, the vertical counterpart of `<hr>`. Handy
  between items in an `bs_hstack()`.

  ```{r}
  bs_hstack(gap = 2, "Left", bs_vr(), "Right")
  ```

- **`bs_visually_hidden()`** — content available to assistive technology but
  hidden from sighted users.

  ```{r}
  bs_visually_hidden("Loading, please wait")
  ```
