---
title: "Forms and inputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Forms and inputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

Form inputs are where `bootstrict` stays closest to Shiny. **Most inputs
delegate to the matching `shiny::*Input()`** and only re-dress the markup as
Bootstrap 5 — so `input$id` and Shiny's own `updateXxx()` work exactly as you
already know. Two native inputs have no Shiny equivalent and get their own
`update_bs_*()` helpers.

## The two families

| Family | Constructors | Reactive value | Update from server |
| --- | --- | --- | --- |
| **Delegating** (wrap Shiny) | text, textarea, numeric, password, select, checkbox, switch, radio, checkbox-group, file | `input$id` | Shiny's `updateXxx()` |
| **Native** (own binding) | `bs_range_input()`, `bs_color_input()`, `bs_date_input()`, `bs_date_range_input()`, `bs_radio_button_input()`, `bs_checkbox_button_input()` | `input$id` | `update_bs_range()`, `update_bs_color()`, `update_bs_date_input()`, `update_bs_toggle_buttons()` |

Every input takes a leading `id`, an optional `label`, and — for delegating
inputs — a `help` string that renders as `.form-text` below the control, plus a
`width`. Where Bootstrap offers sizing, `size = "sm"` or `"lg"` is available.

## Text-like inputs

```{r}
bs_text_input("name", "Your name", placeholder = "Jane Doe")
bs_textarea_input("bio", "Bio", rows = 4)
bs_numeric_input("age", "Age", value = 30, min = 0, max = 120)
bs_password_input("pw", "Password", help = "Must be 8–20 characters.")
```

Signatures share a common shape; `bs_text_input()` is representative:

```{r}
bs_text_input(
  id,
  label = NULL,
  value = "",
  ...,
  placeholder = NULL,
  size = NULL,      # "sm" or "lg"
  help = NULL,      # .form-text below the control
  width = NULL
)
```

Because these delegate to Shiny, update them with Shiny's own updaters:

```{r}
updateTextInput(session, "name", value = "Colin")
```

## Select

`bs_select_input()` renders a **plain Bootstrap `<select class="form-select">`**.
Note this differs from Shiny's `selectInput()`: selectize is off, so there is no
search / tagging box.

```{r}
bs_select_input("fruit", "Fruit", c("Apple", "Pear"))
bs_select_input(
  "sort_by", "Sort by",
  c("Magnitude" = "mag", "Depth" = "depth"),
  selected = "mag"
)
```

`multiple = TRUE` renders a multi-select listbox.

## Checkboxes, switches, radios and groups

A single checkbox, and its switch variant:

```{r}
bs_checkbox_input("agree", "I agree", value = TRUE)
bs_checkbox_input("dark", "Dark mode", switch = TRUE)
bs_switch_input("dark", "Dark mode")            # convenience for switch = TRUE
```

Radio buttons and checkbox groups take `choices` (named or unnamed) and an
optional `selected`. `inline = TRUE` lays the choices out horizontally:

```{r}
bs_radio_input("size", "Size", c("S", "M", "L"), selected = "M")
bs_checkbox_group_input("opts", "Options", c("A", "B", "C"), selected = "A", inline = TRUE)
```

All of the check-style inputs support `reverse = TRUE`, which puts the label
before the control (`.form-check-reverse`).

## Range and colour (native inputs)

These two have no Shiny equivalent, so they ship their own bindings. They still
report `input$id`, but you drive them with the `bootstrict` helpers — Shiny's
`updateSliderInput()` will not reach them.

`bs_range_input()` is a native `<input type="range">` (**not** Shiny's
`sliderInput()` — no ticks, animation or ion.rangeSlider features):

```{r}
bs_range_input("vol", "Volume", value = 50, min = 0, max = 100, step = 1)
```

If `value = NULL`, the browser starts it at the midpoint `(min + max) / 2`.
Update it from the server with `update_bs_range(id, value)`:

```{r}
update_bs_range("vol", 75)
```

`bs_color_input()` is a native colour picker; `value` must be a 6-digit hex
string:

```{r}
bs_color_input("col", "Pick a colour", value = "#0d6efd")
update_bs_color("col", "#198754")
```

## Dates and files

```{r}
bs_date_input("day", "Pick a day", value = "2026-06-26")
bs_date_range_input("range", "Period", start = "2026-01-01", end = "2026-12-31")
bs_file_input("upload", "Upload a file", accept = ".csv")
```

These delegate to `shiny::dateInput()`, `dateRangeInput()` and `fileInput()`
respectively, so update them with the matching Shiny updater.

## Composing inputs

Three layout helpers restyle or group inputs without touching their bindings.

### Input groups

`bs_input_group()` places add-ons and controls on a single line. When you pass a
full `bs_*_input()`, only its bare control is kept (its label and help text are
dropped, since the group is one line):

```{r}
bs_input_group(
  bs_input_group_text("@"),
  bs_text_input("user", placeholder = "Username")
)

bs_input_group(
  bs_input_group_text("$"),
  bs_numeric_input("amount", value = 0),
  bs_input_group_text(".00"),
  size = "lg"
)
```

A handful of inputs whose binding lives on their container —
`bs_date_input()`, `bs_date_range_input()`, `bs_file_input()`,
`bs_radio_input()` and `bs_checkbox_group_input()` — cannot be unwrapped into a
group and will raise a clear error if you try.

### Floating labels

`bs_floating_label()` reshapes an input into the Bootstrap 5 floating-label
style. It **moves** the existing control rather than rebuilding it, so the
binding is preserved. If you do not pass a `label`, the control's existing label
text is reused:

```{r}
bs_floating_label(bs_text_input("email", "Email address"))
bs_floating_label(bs_select_input("fruit", "Fruit", c("Apple", "Pear")))
```

The same container-bound inputs listed above cannot be floated.

### Forms, labels, help and validation

`bs_form()` is a plain `<form>` wrapper for grouping controls; `novalidate =
TRUE` disables the browser's native validation UI:

```{r}
bs_form(
  bs_text_input("email", "Email"),
  bs_button("submit", "Submit", color = "primary")
)
```

The lower-level pieces are available when you hand-build a control:

- `bs_form_label("email", "Email address")` — a `.form-label` tied to a control
  by its `for` attribute.
- `bs_form_text("Must be 8–20 characters.")` — muted helper text (the same
  thing the `help=` argument produces).
- `bs_valid_feedback("Looks good!")` and
  `bs_invalid_feedback("Please choose a username.")` — inline validation
  messages that Bootstrap shows when the control is marked valid/invalid.

## Dates

`bs_date_input()` is a native `<input type="date">` with the `.form-control`
class, which is what the Bootstrap 5.3 forms page shows: the **browser**
supplies the calendar.

It deliberately does not delegate to `shiny::dateInput()`, which loads
`bootstrap-datepicker` — a third-party stylesheet whose calendar markup
(`.datepicker`, `.datepicker-days`, …) appears nowhere in the Bootstrap
documentation and which a designer's SASS sheet cannot reach. The cost is
real: `format`, `language`, `weekstart` and `datesdisabled` are gone, because
the browser owns the presentation and Bootstrap offers no way to ask it for
something else.

```{r}
bs_date_input("day", "Day", value = "2026-06-26", min = "2026-01-01")
bs_date_range_input("period", "Period", start = "2026-01-01")
```

`input$day` is a `Date`, and `NA` while the field is empty. A range is a
length-2 `Date`, so it keeps both positions when only one end is filled in.
Drive them with `update_bs_date_input()` and
`update_bs_date_range_input()` — `shiny::updateDateInput()` will not reach
them. Pass `NA` to clear a field, and `NULL` (the default) to leave it alone.

```{r}
update_bs_date_input("day", value = Sys.Date())
update_bs_date_range_input("period", start = "2026-02-01", end = NA)
```

Bootstrap has no date-range component, so `bs_date_range_input()` is the
assembly its docs prescribe: two fields in an `.input-group` with an
`.input-group-text` between them. Each field carries its own `aria-label`,
since one label cannot name both.

## Toggle buttons

Bootstrap's segmented control is a row of buttons backed by hidden
`.btn-check` inputs. Its markup is not the `.form-check` one with a different
skin -- the input is a *sibling* of its label -- so these are native controls
with their own constructors, driven by `update_bs_toggle_buttons()` rather than
`shiny::updateRadioButtons()`.

```{r}
bs_radio_button_input("size", "Size", c(Small = "s", Large = "l"))
bs_checkbox_button_input("opts", "Options", c("a", "b"))
```

`input$size` is a single string; `input$opts` is a character vector, empty when
nothing is picked. `outline = FALSE`, `color`, `size` and `vertical` cover the
Bootstrap variants.

## Validation

Bootstrap only displays a feedback message when it is a *following sibling* of
the control carrying `.is-valid` / `.is-invalid`. A message placed after a
`bs_*_input()` is a sibling of shiny's input container, not of the control
inside it, so it would never appear. `bs_feedback()` puts it in the right
place:

```{r}
bs_feedback(
  bs_text_input("user", "Username"),
  invalid = "Please choose a username."
)
```

Declare the messages in the UI, then switch the state from the server with
`set_bs_validation()`:

```{r}
observeEvent(input$submit, {
  if (input$user %in% taken) {
    set_bs_validation("user", "invalid", "That name is taken.")
  } else {
    set_bs_validation("user", "valid")
  }
})
```

`set_bs_validation(id, "none")` clears the state. Pass `state =` to
`bs_feedback()` when a control should render already marked.

The bare `bs_valid_feedback()` / `bs_invalid_feedback()` divs remain useful in
a hand-built `<form>`, where you control the sibling order yourself.

## Putting it together

A small settings form, driven from the server:

```{r}
ui <- bs_page(
  bs_container(
    class = "py-4",
    bs_form(
      bs_floating_label(bs_text_input("email", "Email address")),
      bs_range_input("vol", "Volume", value = 50),
      bs_switch_input("notify", "Email me updates", value = TRUE),
      bs_button("save", "Save settings", color = "primary")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$save, {
    str(list(email = input$email, vol = input$vol, notify = input$notify))
    update_bs_range("vol", 50)          # native input → bootstrict helper
    updateTextInput(session, "email", value = "")  # delegating input → Shiny
  })
}

shinyApp(ui, server)
```
