---
title: "Layout and the grid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Layout and the grid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

Layout in `bootstrict` is the Bootstrap 5 layout system, one-to-one: a page, an
optional container, the responsive 12-column grid, and flex stacks for simpler
one-dimensional arrangements.

## Pages

Every app is wrapped in a page constructor. There are three, mirroring `bslib`:

| Constructor | Wraps | Use for |
| --- | --- | --- |
| `bs_page()` | `bslib::page()` | the default fixed-width page |
| `bs_page_fluid()` | `bslib::page_fluid()` | a full-width page |
| `bs_page_fillable()` | `bslib::page_fillable()` | a page whose content fills the viewport height |

All three share the same signature:

```{r}
bs_page(
  ...,                          # page content (and named HTML attributes)
  title = NULL,                 # browser tab / document title
  theme = bootstrict_theme(),   # a bootstrict_theme() / bslib::bs_theme()
  color_mode = NULL,            # "light" or "dark" — sets data-bs-theme
  lang = "en"
)
```

A page wires in the `bootstrict` dependency and the theme, so it is the only
place you configure them:

```{r}
bs_page(
  title = "My app",
  theme = bootstrict_theme(primary = "#ff6600"),
  color_mode = "light",
  bs_container(bs_card(bs_card_body("Hello")))
)
```

See [Theming](theming.html) for `theme` and `color_mode`.

## Containers

`bs_container()` centres and horizontally pads its content. By default it is
fixed-width with a responsive max-width; `fluid = TRUE` makes it span the full
width, and `breakpoint=` makes it fluid *until* a chosen breakpoint.

```{r}
bs_container(
  ...,
  fluid = FALSE,       # TRUE for a full-width .container-fluid
  breakpoint = NULL,   # "sm"/"md"/"lg"/"xl"/"xxl" for .container-{bp}
  class = NULL
)
```

```{r}
bs_container(bs_row(bs_col("a"), bs_col("b")))     # fixed width
bs_container(fluid = TRUE, "full-width content")   # 100% wide
bs_container(breakpoint = "lg", "fluid below lg")  # full width until lg
```

## The grid: rows and columns

Bootstrap's grid is a set of rows, each split into up to 12 columns.

### `bs_row()`

```{r}
bs_row(
  ...,             # bs_col() children (and named attributes)
  cols = NULL,     # equal-width columns per row: integer 1-6, or "auto"
  gutters = NULL,  # gutter size 0-5 (both axes)
  gx = NULL,       # horizontal gutter 0-5
  gy = NULL,       # vertical gutter 0-5
  justify = NULL,  # "start"/"center"/"end"/"around"/"between"/"evenly"
  align = NULL,    # cross-axis: "start"/"center"/"end"
  class = NULL
)
```

### `bs_col()`

```{r}
bs_col(
  ...,               # column content
  width = NULL,      # base span: 1-12, "auto", or TRUE (equal-width)
  sm = NULL,         # span from the sm breakpoint up
  md = NULL,         # ... md
  lg = NULL,         # ... lg
  xl = NULL,         # ... xl
  xxl = NULL,        # ... xxl
  offset = NULL,     # push right by 0-11 columns
  order = NULL,      # 0-5, "first" or "last"
  align_self = NULL, # "start"/"center"/"end"
  class = NULL
)
```

A column with no `width` is a bare `.col` and shares the row's space equally
with its siblings:

```{r}
bs_row(bs_col("a"), bs_col("b"))   # two equal columns
```

### Responsive spans

The breakpoint arguments (`width`, `sm`, `md`, …) let one column change width as
the viewport grows. A common pattern — full width on phones, a third of the row
from `md` up:

```{r}
bs_container(
  bs_row(
    class = "g-3",
    bs_col(md = 4, bs_card(bs_card_body("A"))),
    bs_col(md = 4, bs_card(bs_card_body("B"))),
    bs_col(md = 4, bs_card(bs_card_body("C")))
  )
)
```

Here `width` is unset, so each column is full-width (`.col-12`) below `md` and
becomes `.col-md-4` from `md` up.

### Gutters, alignment and order

`gutters` (or the per-axis `gx` / `gy`) control the spacing between columns;
`justify` and `align` position them within the row; `offset` and `order` push
and reorder individual columns:

```{r}
bs_row(
  gutters = 3,
  justify = "between",
  bs_col(width = 4, "left"),
  bs_col(width = 4, offset = 0, order = "last", "shown last")
)
```

Several grid arguments also accept a **named per-breakpoint list** for
responsive behaviour — for instance a wider gutter only on large screens, or a
different equal-column count per breakpoint:

```{r}
bs_row(cols = list(sm = 1, md = 2, lg = 3), bs_col("..."), bs_col("..."))
bs_col(offset = list(md = 2), "offset by two columns from md up")
```

## Stacks

For simple one-dimensional layouts, flex **stacks** are lighter than the grid.
`bs_hstack()` lays children out in a row, `bs_vstack()` in a column; `gap=`
(0-5) sets the spacing between them.

```{r}
bs_hstack(
  bs_button(label = "A"),
  bs_button(label = "B"),
  gap = 2
)

bs_vstack(
  bs_alert("one"),
  bs_alert("two"),
  gap = 3
)
```

Because stacks are plain flex containers, they combine naturally with Bootstrap
utility classes passed through `class` — for example to right-align a toolbar:

```{r}
bs_hstack(
  gap = 3,
  class = "justify-content-end align-items-center",
  bs_switch_input("dark", "Dark mode", width = "auto"),
  bs_button("open", "Filters", color = "primary")
)
```
