---
title: "Overlays and server-driven interactivity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overlays and server-driven interactivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

This article covers the overlay widgets — modal, offcanvas, toast — plus
tooltips, popovers and scrollspy, and pulls together the one pattern that runs
through the whole package: interactive widgets report their state as `input$id`
and are controlled from the server with `id`-first helpers.

## The overlay model: declare in the UI, open by id

This is the biggest deliberate difference from Shiny. Shiny builds overlays on
the server (`showModal(modalDialog(...))`). `bootstrict` follows the native
Bootstrap pattern: **declare the overlay once in the UI with an `id`, then open
or close it by id from the server.**

Two consequences follow:

- **Place overlays at the top level of the page** — a direct child of
  `bs_page()` / `bs_container()`, *not* nested inside a `bs_card()` or other
  positioned element, which Bootstrap can clip or mis-position.
- **Every overlay reports its open state as `input$id`** (`TRUE` when shown) —
  something Shiny modals do not do.

## Modal

Declare a `bs_modal()` with an `id`; open it with `show_bs_modal()`.

```{r}
ui <- bs_page(
  bs_container(
    bs_button("open", "Open modal", color = "primary")
  ),
  # top level of the page, not inside the container's card
  bs_modal(
    "info",
    "Modal body text.",
    title = "Heads up",
    footer = bs_button("ok", "OK", color = "primary")
  )
)

server <- function(input, output, session) {
  observeEvent(input$open, show_bs_modal("info"))
  observeEvent(input$info, message("modal open state: ", input$info))
}

shinyApp(ui, server)
```

The server helpers are `show_bs_modal()`, `hide_bs_modal()` and
`toggle_bs_modal()` — each `id`-first, session-last:

```{r}
show_bs_modal("info")
hide_bs_modal("info")
```

Options: `size` (`"sm"`/`"lg"`/`"xl"`), `centered`, `scrollable`,
`fullscreen` (`TRUE`, or a breakpoint for fullscreen-below), `backdrop`
(`TRUE` / `"static"` / `FALSE`) and `keyboard` (Escape to close). The pieces
`bs_modal_header()`, `bs_modal_title()`, `bs_modal_body()` and
`bs_modal_footer()` are there when you want to hand-build the dialog.

### Dynamic content

An overlay is declared once, but its content need not be fixed: put a
`uiOutput()` in the body and render into it. This is how you show a record the
user just picked, without rebuilding the modal.

```{r}
ui <- bs_page(
  bs_container(
    bs_button("show", "Show details"),
    bs_modal("details", uiOutput("details_body"), title = "Details")
  )
)

server <- function(input, output, session) {
  output$details_body <- renderUI({
    bs_table(head(mtcars, input$show %% 5 + 1))
  })
  observeEvent(input$show, show_bs_modal("details"))
}
```

Re-rendering while the overlay is open is safe: the widget is torn down after
its hide transition, so the page is never left scroll-locked.

### Opening without a server round trip

`bs_modal_trigger()` opens a modal declaratively from the UI — no `observeEvent`
needed:

```{r}
bs_modal_trigger("info", "Open modal", color = "primary")
```

## Offcanvas (drawer)

An offcanvas is a panel that slides in from an edge — Shiny has no equivalent.
Same model as the modal: declare it, open it by id.

```{r}
ui <- bs_page(
  bs_container(
    bs_offcanvas_trigger("filters", "Filters", class = "btn-primary")
  ),
  bs_offcanvas(
    "filters",
    title = "Filter events",
    placement = "end",              # "start"/"end"/"top"/"bottom"
    bs_range_input("mag", "Minimum magnitude", value = 4, min = 4, max = 6.5),
    bs_button("reset", "Reset", color = "secondary", outline = TRUE)
  )
)

server <- function(input, output, session) {
  observeEvent(input$reset, {
    update_bs_range("mag", 4)
    hide_bs_offcanvas("filters")
  })
}

shinyApp(ui, server)
```

Server helpers: `show_bs_offcanvas()`, `hide_bs_offcanvas()`,
`toggle_bs_offcanvas()`, plus the UI `bs_offcanvas_trigger()`. `backdrop`
behaves as for the modal; `responsive = "lg"` shows the panel inline from that
breakpoint up and only turns it into a drawer below it.

## Toasts

A toast is a small, transient notification. There are two ways to use them.

**Declared toast.** Declare a `bs_toast()` (usually inside a
`bs_toast_container()` that positions it) and show it by id:

```{r}
ui <- bs_page(
  bs_button("save", "Save"),
  bs_toast_container(
    bs_toast("saved", "Your changes were saved.", title = "Done"),
    placement = "top-end"
  )
)

server <- function(input, output, session) {
  observeEvent(input$save, show_bs_toast("saved"))
}
```

**Notification toast.** `bs_notify_toast()` is the one server-built widget — the
closest thing to `shiny::showNotification()`. It builds *and* shows a transient
toast in one call, creating a container on demand and cleaning itself up
afterwards. Note the argument order: `body` first, `session` last.

```{r}
bs_notify_toast(
  "Showing the full catalogue again.",
  title = "Filters reset",
  color = "success"
)
```

`body` and `title` are plain text only. `color` is a theme colour (`.text-bg-*`),
`autohide`/`delay` control the auto-dismiss, and `placement` uses the same nine
keywords as `bs_toast_container()` (`"top-end"`, `"bottom-center"`, …).

## Tooltips and popovers

These **decorate a tag you already have** — they are pipe-friendly wrappers, not
constructors. Bootstrap does not auto-initialise them, but `bootstrict` does.
They are UI-only: there is no server-side toggle.

```{r}
bs_tooltip(bs_button(label = "Hover me"), "Tooltip text", placement = "top")

bs_popover(
  bs_button(label = "Click me"),
  "Popover body content",
  title = "Heads up",
  placement = "right"
)
```

`bs_tooltip()` takes `title`; `bs_popover()` takes `content` (and an optional
`title`). Both accept `placement` (`"top"`/`"right"`/`"bottom"`/`"left"`),
`trigger` and `html = TRUE` to allow HTML content.

## Scrollspy

`bs_scrollspy()` wraps a scrollable region and updates a nav (identified by
`target`) as the user scrolls. It needs an `id` to initialise (one is
auto-generated if you omit it), including inside `renderUI()`; the active nav
link's href is reported as `input$id`.

```{r}
bs_scrollspy(
  "section-nav",
  tags$h4("Section one"), tags$p("..."),
  tags$h4("Section two"), tags$p("..."),
  offset = 100
)
```

## Inside a Shiny module

The two halves are not symmetric, and it is the one place bootstrict asks you
to think.

**Server helpers namespace for you.** `show_bs_modal("info")` inside a
`moduleServer()` resolves to the module's `info`, because the helper reads the
session's namespace. Never wrap them in `ns()`.

**UI triggers do not.** `bs_modal_trigger()`, `bs_offcanvas_trigger()` and
`bs_collapse_trigger()` take the id of a *target element*, which is a plain
string at UI build time with no session to consult. Wrap those in `ns()`
yourself, exactly as you would a `tabPanel` id.

```{r}
panel_ui <- function(id) {
  ns <- NS(id)
  bs_container(
    # The widget's own id: namespaced, like any input.
    bs_modal(ns("info"), "Body", title = "Heads up"),
    # A trigger points at that element, so it needs the same treatment.
    bs_modal_trigger(ns("info"), "Open"),
    bs_button(ns("open"), "Open from the server")
  )
}

panel_server <- function(id) {
  moduleServer(id, function(input, output, session) {
    # No ns() here: the helper takes it from the session.
    observeEvent(input$open, show_bs_modal("info"))
  })
}
```

The same rule covers `bs_collapse_trigger()` pointing at a `bs_collapse()`, and
`bs_scrollspy(target =)` pointing at a nav.

## The pattern, in one place

Every interactive widget in `bootstrict` follows the same two-way contract:

| Widget | Reads back as `input$id` | Server control |
| --- | --- | --- |
| `bs_tabset()` | active panel value | `update_bs_tabset()` |
| `bs_accordion()` | open panel value(s) | `update_bs_accordion()` |
| `bs_carousel()` | active slide index (0-based) | `update_bs_carousel()` |
| `bs_collapse()` | shown/hidden (`TRUE`/`FALSE`) | `update_bs_collapse()` |
| `bs_list_group(id=)` | selected item value | `update_bs_list_group()` |
| `bs_progress_bar(id=)` | *(display only)* | `update_bs_progress()` |
| `bs_range_input()` | value | `update_bs_range()` |
| `bs_color_input()` | value | `update_bs_color()` |
| `bs_modal()` | open (`TRUE`/`FALSE`) | `show_/hide_/toggle_bs_modal()` |
| `bs_offcanvas()` | open (`TRUE`/`FALSE`) | `show_/hide_/toggle_bs_offcanvas()` |
| `bs_toast()` | visible (`TRUE`/`FALSE`) | `show_/hide_bs_toast()` |
| page colour mode | — | `set_bs_color_mode()` |

Every one of these helpers takes the **`id` first** and the **`session` last and
optional** (it defaults to the current reactive domain), and ids are namespaced
automatically inside Shiny modules — so the same call works at the top level and
inside a module with no changes:

```{r}
update_bs_tabset("tabs", selected = "profile")   # no session argument needed
show_bs_modal("info")
set_bs_color_mode("dark")
```
