---
title: "Theming and the designer hand-off"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Theming and the designer hand-off}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(shiny)
library(bootstrict)
```

`bootstrict` vendors Bootstrap 5.3 itself — `bootstrap_version()` reports the
exact release — and compiles it with [`sass`](https://rstudio.github.io/sass/),
so the stylesheet reaching the browser is pinned by this package rather than by
whichever theming package happens to be installed. Because the markup
`bootstrict` emits is *exactly* the markup Bootstrap 5.3 documents, a
designer's SASS variables restyle every widget with no per-component work. This
article covers the whole theming story: building a theme, the designer hand-off
via a `_variables.scss` sheet, and Bootstrap 5.3 colour modes.

## The motivating workflow

The workflow `bootstrict` is built around:

1. A designer works in Figma, staying strictly within the Bootstrap 5.3 docs.
2. They export a `_variables.scss` sheet — a list of `$name: value;` overrides.
3. You receive the mockup and the sheet, and wire the sheet into the app. That
   is the *entire* theming step.

```{r}
ui <- bs_page(
  theme = bootstrict_theme(variables = "_variables.scss"),
  bs_container(
    bs_card(bs_card_body("The whole visual identity comes from the sheet."))
  )
)
```

## `bootstrict_theme()`

`bootstrict_theme()` collects the SASS variable overrides to compile the
vendored Bootstrap tree with. It merges, in order of increasing priority:

- a designer's exported SASS variable sheet (`variables=`),
- inline variables passed through `...`.

```{r}
bootstrict_theme(
  ...,              # Sass variables, e.g. primary = "#ff6600"
  variables = NULL  # path to a .scss sheet, or a named list
)
```

Inline `...` variables **win over the file** on a name clash — handy for
overriding a single token without editing the designer's sheet:

```{r}
bootstrict_theme(
  variables = "_variables.scss",
  primary = "#ff6600"
)
```

Any Bootstrap SASS variable works here, named without its `$`. Note that some
of those names are not valid R names, so quote them:

```{r}
bootstrict_theme(primary = "#ff6600", "font-size-base" = "1rem")
```

Pass the resulting theme to any page constructor:

```{r}
bs_page(theme = bootstrict_theme(primary = "#0d6efd"), ...)
```

## The `_variables.scss` sheet

A `_variables.scss` sheet is plain SCSS: one `$name: value;` declaration per
line. It is exactly what a designer exports, and it drops in unchanged.

```scss
// _variables.scss
$primary:       #ff6600;
$secondary:     #6c757d;
$border-radius: 0.5rem;
$font-family-base: "Inter", sans-serif;
```

### Inspecting a sheet with `parse_scss_variables()`

`parse_scss_variables()` turns a sheet's top-level declarations into a named
list. `bootstrict_theme(variables=)` calls it for you, but it is
useful on its own — for example to echo the active theme back in an app.

```{r}
tmp <- tempfile(fileext = ".scss")
writeLines(
  c("$primary: #ff6600;", "$border-radius: 0.5rem !default;"),
  tmp
)
parse_scss_variables(tmp)
#> $primary
#> [1] "#ff6600"
#>
#> $`border-radius`
#> [1] "0.5rem"
```

It strips `!default` / `!global` flags and comments. Note the parser reads the
**SCSS** syntax (`$name: value;`, with semicolons) — the indented `.sass`
syntax has no semicolons and cannot be parsed.

## Colour modes (Bootstrap 5.3)

Bootstrap 5.3 introduced native light/dark **colour modes** driven by the
`data-bs-theme` attribute. `bootstrict` exposes them at three levels.

### Set the initial mode on the page

```{r}
bs_page(color_mode = "dark", ...)
```

### Flip the mode from the server

`set_bs_color_mode()` toggles the whole page between `"light"` and `"dark"` at
runtime:

```{r}
server <- function(input, output, session) {
  observeEvent(input$dark_switch, {
    set_bs_color_mode(if (isTRUE(input$dark_switch)) "dark" else "light")
  })
}
```

A complete light/dark toggle is just a switch wired to `set_bs_color_mode()`:

```{r}
ui <- bs_page(
  color_mode = "light",
  bs_container(
    bs_switch_input("dark", "Dark mode"),
    bs_card(bs_card_body("This surface follows the colour mode."))
  )
)

server <- function(input, output, session) {
  observeEvent(input$dark, {
    set_bs_color_mode(if (isTRUE(input$dark)) "dark" else "light")
  })
}

shinyApp(ui, server)
```

### Following the operating system

`color_mode = "auto"` resolves the mode from the OS
(`prefers-color-scheme`) before the page paints, so there is no flash of the
wrong theme. A mode the user later chooses is remembered in the browser and
wins over the initial value on the next visit; `set_bs_color_mode("auto")`
forgets it and hands control back to the OS.

Whatever the preference, the mode actually in force is reported as
`input$bootstrict_color_mode`, always `"light"` or `"dark"` -- never
`"auto"` -- so the server can render to match.

```{r}
ui <- bs_page(
  color_mode = "auto",
  bs_container(
    bs_radio_button_input(
      "mode",
      "Appearance",
      c(Light = "light", Dark = "dark", Auto = "auto"),
      selected = "auto"
    ),
    plotOutput("plot")
  )
)

server <- function(input, output, session) {
  observeEvent(input$mode, set_bs_color_mode(input$mode))

  output$plot <- renderPlot({
    # Follows the resolved mode, not the preference.
    bg <- if (identical(input$bootstrict_color_mode, "dark")) "black" else "white"
    plot(1, bg = bg)
  })
}
```

An app that never mentions `color_mode` is left alone: it keeps Bootstrap's
default rather than quietly following the OS.

### Scope a mode to one component

Many components accept a component-level `dark = TRUE` (or `theme = "dark"`)
argument, which emits `data-bs-theme` on just that element — the 5.3 idiom that
replaces the deprecated `.navbar-dark` / `.dropdown-menu-dark` /
`.carousel-dark` classes. For example a dark navbar on a light page:

```{r}
bs_navbar(brand = bs_navbar_brand("Acme"), bg = "primary", theme = "dark")
```

## Utilities

Bootstrap's [utility classes](https://getbootstrap.com/docs/5.3/utilities/) --
spacing, colour, display, flex, borders, text -- are the other half of what a
designer works with, and bootstrict deliberately exposes none of them as
arguments. Every constructor takes a trailing `class` and forwards named `...`
as HTML attributes, so a utility is written as itself:

```{r}
bs_card(
  bs_card_body("Padded, centred, muted."),
  class = "p-4 text-center text-body-secondary"
)
bs_row(
  bs_col("Left", md = 6),
  bs_col("Right", md = 6),
  class = "gy-3 align-items-center"
)
```

That is the whole story: there is no `padding =` argument to learn, and the
mockup's class list transfers verbatim. It also means a utility that Bootstrap
adds in a future release works here on the day it ships.

The
[Utility API](https://getbootstrap.com/docs/5.3/utilities/api/) belongs to the
theme rather than to the markup: extend or disable utilities from the SASS
sheet, alongside the variables.

```scss
$utilities: map-merge(
  $utilities,
  ("cursor": (property: cursor, class: cursor, values: pointer grab))
);
```

`parse_scss_variables()` reads that map like any other declaration, so it
reaches `bootstrict_theme()` with the rest of the sheet.

## The golem hook

If you scaffold Shiny apps with [`golem`](https://thinkr-open.github.io/golem/),
`use_bootstrict_golem` is a `project_hook` that turns a fresh golem skeleton
into a minimal `bootstrict` app: it rewrites `R/app_ui.R` to use `bs_page()` /
`bs_container()` and creates `inst/app/www/_variables.scss` for you.

```{r}
golem::create_golem(
  "my.app",
  project_hook = bootstrict::use_bootstrict_golem
)
```

## Attaching the dependency by hand

Page constructors (`bs_page()` and friends) wire in the `bootstrict`
dependency automatically, and every top-level widget carries it too. If you
build a UI without a `bs_page()` wrapper — for example inside a larger app
shell — add the dependency once with `use_bootstrict()`:

```{r}
tagList(
  use_bootstrict(),
  bs_card(bs_card_body("..."))
)
```
